### 2017-03-31 manuscript/fig_aggregation.tex
Move paper to git

\begin{figure}
  \def\aggfg#1#2{%
    \rput[b](#1){\newmatrix{#2}{16pt}{\coordstep}}
    \leftlabel{#2}{$\tilde{a}_f$}
    \refs{#2}{\coordstep}{1}
         {\scriptsize  \canonicalref{#2}{\coordstep}}
     \plotmatrix[\coordstep]{\plotframe{upr}}{naf}{2}{1}
  }

  \begin{center}
    \begin{pspicture}(0,-1)(16.5,10)\showgrid
      \let\showgrid\relax
      \rput[tl](0,10){\input{figures/fig4a-unitprocess.fig}}
      \rput[tl](2.7,10){\input{figures/fig4b-foreground.fig}}
      \rput[tl](5.4,10){\input{figures/fig4c-agg-foreground.fig}}
      \rput[tl](8.1,10){\input{figures/fig4d-partial-bg-agg.fig}}
      \rput[tl](11.6,10){\input{figures/fig4e-full-agg.fig}}
      \rput[tl](15.1,10){\input{figures/fig4f-lci.fig}}
    \end{pspicture}
    \caption{Publication structures for different forms of foreground aggregation.}
    \label{fig:aggregation}
  \end{center}
\end{figure}