### 2017-03-31 manuscript/figures/potato-organic.tex

market for potato, organic [GLO]

{\scriptsize\sffamily
\begin{tabularx}{\textwidth}{|>{\hangindent=2ex}X|c@{~}c@{~}c@{~}c@{~}c@{~}c@{~}c@{~}c@{~}c@{~}|c|}
\hline
(node) Foreground flow \rule[-3pt]{0pt}{12pt} & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \\ 
\hline
(0) potato, organic (kg) [GLO] & \refbox  &  &  &  &  &  &  &  &  & \\ 
(1) potato, organic (kg) [CH] & 0.024 & \refbox  &  &  &  &  &  &  &  & \\ 
(2) potato, organic (kg) [RoW] & 0.976 &  & \refbox  &  &  &  &  &  &  & \\ 
(3) potato seed, organic, at farm (kg) [RoW] &  &  &  & \refbox  &  &  &  & 0.976 &  & \\ 
(4) potato seed, organic, for setting (kg) [CH] &  &  &  &  & \refbox  &  &  &  & 0.024 & \\ 
(5) potato seed, organic, for setting (kg) [RoW] &  &  &  &  &  & \refbox  &  &  & 0.976 & \\ 
(6) potato seed, organic, at farm (kg) [CH] &  &  &  &  &  &  & \refbox  & 0.024 &  & \\ 
(7) potato seed, organic, at farm (kg) [GLO] &  &  &  &  &    1 &    1 &  & \refbox  &  & \\ 
(8) potato seed, organic, for setting (kg) [GLO] &  & 0.11 & 0.11 & 0.16 &  &  & 0.16 &  & \refbox  & \\ 
\hline
Foreground Node Weights $\tilde{x}$ &    1 & 0.024 & 0.976 & 0.128 & 0.00315 & 0.128 & 0.00315 & 0.131 & 0.131& \\ 
\hline
\hline
Background Dependencies \rule[-3pt]{0pt}{12pt} & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & $\tilde{a_d}$\\ 
\hline
transport, freight train (metric ton*km) [GLO] & \dependency &  &  &  &  &  &  & \dependency &  & 0.138\\ 
irrigation (m3) [CH] &  & \dependency &  &  &  &  & \dependency &  &  & 0.000426\\ 
tillage, hoeing and earthing-up, potatoes (ha) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 0.000101\\ 
potato planting (ha) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 5.04e-05\\ 
solid manure loading and spreading, by hydraulic loader and spreader (kg) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 0.717\\ 
electricity, low voltage (kWh) [CH] &  &  &  &  & \dependency &  &  &  &  & 0.00023\\ 
potato grading (kg) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  &  1.13\\ 
transport, freight, light commercial vehicle (metric ton*km) [GLO] & \dependency &  &  &  &  &  &  & \dependency &  & 0.0258\\ 
tillage, harrowing, by spring tine harrow (ha) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 5.04e-05\\ 
tillage, ploughing (ha) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 5.04e-05\\ 
electricity, low voltage (kWh) [GLO] &  &  &  &  &  & \dependency &  &  &  & 0.00933\\ 
transport, freight, lorry, unspecified (metric ton*km) [GLO] & \dependency &  &  &  &  &  &  & \dependency &  & 0.404\\ 
green manure, organic, until March (ha) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 5.04e-05\\ 
building, multi-storey (m3) [GLO] &  &  &  &  & \dependency & \dependency &  &  &  & 5.24e-06\\ 
liquid manure spreading, by vacuum tanker (m3) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 0.000565\\ 
application of plant protection product, by field sprayer (ha) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 0.000222\\ 
transport, freight, sea, transoceanic ship (metric ton*km) [GLO] & \dependency &  &  &  &  &  &  & \dependency &  & 0.525\\ 
potato haulm cutting (ha) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 5.04e-05\\ 
transport, tractor and trailer, agricultural (metric ton*km) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 0.00113\\ 
harvesting, by complete harvester, potatoes (ha) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 5.04e-05\\ 
tillage, currying, by weeder (ha) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 0.000101\\ 
transport, freight, inland waterways, barge (metric ton*km) [GLO] & \dependency &  &  &  &  &  &  & \dependency &  & 0.0857\\ 
copper oxide (kg) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 0.000107\\ 
tillage, harrowing, by rotary harrow (ha) [GLO] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 5.04e-05\\ 
irrigation (m3) [GLO] &  &  & \dependency & \dependency &  &  &  &  &  & 0.0173\\ 
\hline
Foreground Emissions \rule[-3pt]{0pt}{12pt} & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & $\tilde{b_f}$\\ 
\hline
Input: Occupation, construction site [natural resource, land] &  &  &  &  & \dependency & \dependency &  &  &  & 2.1e-06\\ 
Input: Transformation, from unspecified [natural resource, land] &  &  &  &  & \dependency & \dependency &  &  &  & 1.05e-06\\ 
Input: Transformation, to industrial area [natural resource, land] &  &  &  &  & \dependency & \dependency &  &  &  & 1.05e-06\\ 
Input: Energy, gross calorific value, in biomass [natural resource, biotic] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  &  3.87\\ 
Output: Phosphate [water, ground-] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 6.25e-06\\ 
Output: Cadmium, ion [water, ground-] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 1.14e-09\\ 
Output: Zinc, ion [water, ground-] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 1.35e-06\\ 
Output: Nitrate [water, ground-] &  & \dependency & \dependency & \dependency &  &  & \dependency &  &  & 0.00696\\ 
$\ldots$ (31 rows omitted) &  &  &  &  &  &  &  &  & \\ 
\hline
\end{tabularx}
}