### 2017-03-31 manuscript/publication-acknowledgments.tex

  \section*{Acknowledgments}

  This work draws on the efforts of the ``Inventory Model Description and Revision'' discussion group of SETAC North America, whose members include Ben Mourad Amor (University of Sherbrooke), Bill Bernstein (NIST), Paula Bernstein (PR\'e Sustainability), Marcos Esterman (Rochester Institute of Technology), David Evers (Hexion), Karl Haapala (Oregon State University), Troy Hawkins (Eastern Research Group), Wesley Ingwersen (US EPA), Christoph Koffler (Thinkstep), Lise Laurin (EarthShift Global), Antonino Marvuglia (Luxembourg Institute of Science and Technology), David Meyer (US EPA), KC Morris (NIST), Christopher Mutel (Paul Scherrer Institut), Tomas Navarrete (Luxembourg Institute of Science and Technology), Devarajan Ramanujan (Massachusetts Institute of Technology), Barclay Satterfield (Eastman Chemical), and the author.  This work was supported by the US National Science Foundation (CCF-1442966).