### 2017-03-31 manuscript/publication-si-openaccess.tex
Jupyter notebook for USLCI

\section{Open Access to Paper Materials}
The software used to generate the examples published in this paper is available online.  Please consult the GitHub repository found at \url{http://github.com/bkuczenski/Publication-JIE}.  The python code used to produce the examples is found in the `python` directory, as a set of Jupyter notebooks.  Because the Ecoinvent database cannot be distributed due to licensing restrictions, only the US LCI example works ``out of the box.''

Please contact the author with any questions about the code, via the GitHub page or via email at \texttt{bkuczenski@ucsb.edu}.