### 2017-04-01 manuscript/fig_aggregation.tex
Fig mastering

\begin{figure}
  \begin{center}
    \input{fig4-aggregation.fig}
    \caption{Publication structures for different forms of foreground aggregation.}
    \label{fig:aggregation}
  \end{center}
\end{figure}