### 2017-09-19 figures/aluminum-secondary.tex
Move figures folder

Aluminum, secondary, ingot, from automotive scrap, at plant [RNA]

{\scriptsize\sffamily
\begin{tabularx}{\textwidth}{|>{\hangindent=2ex}X|c@{~}c@{~}c@{~}c@{~}|c|}
\hline
\everypar{\hangindent1em \hangafter1}
(node) Foreground flow \rule[-3pt]{0pt}{12pt} & 0 & 1 & 2 & 3 & \\ 
\hline
(0) Aluminum, secondary, ingot, from automotive scrap, at plant (kg) [RNA] & \refbox  &  &  &  & \\ 
(1) Aluminum recovery, transport, to plant (kg) [RNA] & 1.03 & \refbox  &  &  & \\ 
(2) Quicklime, at plant (kg) [RNA] & 2.35e-05 &  & \refbox  &  & \\ 
(3) Limestone, at mine (kg) [RNA] &  &  & 1.87 & \refbox  & \\ 
\hline
Foreground Node Weights $\tilde{x}$ &    1 & 1.03 & 2.35e-05 & 4.39e-05& \\ 
\hline
Input: CUTOFF Disposal, solid waste, unspecified, to sanitary landfill [CUTOFF Flows] & \dependency &  & \dependency &  & 0.0842\\ 
Input: CUTOFF Filter media, at plant [CUTOFF Flows] & \dependency &  &  &  & 4.79e-05\\ 
Input: CUTOFF Lube oil, at plant [CUTOFF Flows] & \dependency &  &  &  & 8.75e-07\\ 
Input: CUTOFF Treatment gases, unspecified, at plant [CUTOFF Flows] & \dependency &  &  &  & 0.003\\ 
Input: CUTOFF Alloying additives, at plant [CUTOFF Flows] & \dependency &  &  &  & 0.021\\ 
Input: CUTOFF Chemicals, unspecified, used for wastewater treatment [CUTOFF Flows] & \dependency &  &  &  & 0.008\\ 
Input: CUTOFF Grain refiners, at plant [CUTOFF Flows] & \dependency &  &  &  & 0.0003\\ 
Input: CUTOFF Treatment salts, unspecified, at plant [CUTOFF Flows] & \dependency &  &  &  & 0.009\\ 
Input: CUTOFF Packaging, unspecified, at plant [CUTOFF Flows] & \dependency &  &  &  & 1.67e-05\\ 
Input: CUTOFF Aluminum scrap, automotive [CUTOFF Flows] & \dependency &  &  &  &  1.03\\ 
\hline
\hline
Background Dependencies \rule[-3pt]{0pt}{12pt} & 0 & 1 & 2 & 3 & $\tilde{a_d}$\\ 
\hline
Diesel, combusted in industrial boiler (l) [RNA] &  &  & \dependency & \dependency & 4.79e-08\\ 
Transport, train, diesel powered (t*km) [RNA] &  & \dependency & \dependency &  & 0.0415\\ 
Electricity, at grid, US, 2000 (kWh) [RNA] & \dependency &  & \dependency & \dependency & 0.668\\ 
Transport, combination truck, diesel powered (t*km) [RNA] &  & \dependency & \dependency &  & 0.374\\ 
Natural gas, combusted in industrial boiler (m3) [RNA] & \dependency &  & \dependency & \dependency & 0.223\\ 
Bituminous coal, combusted in industrial boiler (kg) [RNA] &  &  & \dependency & \dependency & 4.04e-06\\ 
Gasoline, combusted in equipment (l) [RNA] &  &  &  & \dependency & 2.25e-09\\ 
Liquefied petroleum gas, combusted in industrial boiler (l) [RNA] &  &  & \dependency &  & 7.57e-10\\ 
Transport, barge, average fuel mix (t*km) [RNA] &  &  & \dependency &  & 5.66e-07\\ 
\hline
Foreground Emissions \rule[-3pt]{0pt}{12pt} & 0 & 1 & 2 & 3 & $\tilde{b_f}$\\ 
\hline
Output: Lead [air, unspecified] & \dependency &  &  &  & 2.16e-07\\ 
Output: Suspended solids, unspecified [water, unspecified] & \dependency &  &  &  & 2.56e-05\\ 
Output: Particulates, unspecified [air, unspecified] & \dependency &  & \dependency & \dependency & 1.78e-07\\ 
Output: NMVOC, non-methane volatile organic compounds [air, unspecified] & \dependency &  &  &  & 4.7e-05\\ 
Output: BOD5, Biological Oxygen Demand [water, unspecified] & \dependency &  &  &  & 1.35e-09\\ 
Output: COD, Chemical Oxygen Demand [water, unspecified] & \dependency &  &  &  & 8.3e-07\\ 
Output: Sulfur dioxide  [air, unspecified] &  &  & \dependency &  & 3.52e-09\\ 
Output: Dissolved solids [water, unspecified] & \dependency &  &  &  & 2.38e-08\\ 
Output: carbon dioxide [air, unspecified] &  &  & \dependency &  & 1.8e-05\\ 
Input: Limestone [resource, ground-] &  &  &  & \dependency & 4.39e-05\\ 
Output: Heavy metals, unspecified [water, unspecified] & \dependency &  &  &  & 2.25e-05\\ 
Output: Organic substances, unspecified [water, unspecified] & \dependency &  &  &  & 1.29e-06\\ 
Output: Acids, unspecified [air, unspecified] & \dependency &  &  &  & 4.31e-05\\ 
\hline
\end{tabularx}
}