### 2017-10-17 aggregation/agg-abstract.tex

\begin{abstract}
  Many of the challenges facing %interpretation and
  knowledge synthesis from life cycle assessment (LCA) studies stem from the inability of study authors and readers to formally agree on the structure and content of life cycle inventory models.  This article presents a framework for computing aggregation results in LCA $\ldots$
\end{abstract}