### 2017-10-24 aggregation/agg-review.tex

\subsection{Critical Review of PSMs}

When a study result is published in an aggregated form, it is the responsibility of the critical reviewer to validate the aggregation result.