### 2017-10-30 aggregation/agg-semantic.tex

\subsection{Avoiding Ambiguous Disclosures}

either the column of $B_x$, or the scalar value that corresponds to the dot product of this column with $\mathbf{e}$

Multiple background databases

\subsection{Stable Semantic References}