### 2017-11-02 aggregation/agg-JIE.tex
Finish discussion?

\documentclass[12pt]{article}

\input{formatting}

\usepackage{tabularx}
\usepackage[round]{natbib}
\bibliographystyle{authordate_bk}

\renewcommand{\topfraction}{0.7}

%\thispagestyle{empty}

\input{psstyle}
\input{../manuscript/draw-matrices}

\title{Disclosure of Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy}
\author{Brandon Kuczenski\\
Institute for Social, Behavioral, and Economic Research\\
University of California, Santa Barbara}
\date{\mydate\today}

%\usepackage{lineno}
\let\showgrid\relax
\let\oldtimes\times
\let\times\cdot

\def\mainpaper{

  \maketitle
  \emph{For submission to the Journal of Industrial Ecology; October 2017}

  \input{agg-abstract}

  \doublespacing

  {\red http://pubs.acs.org/doi/pdf/10.1021/acs.est.5b01558}

  \input{agg-intro}  % motivations

\section{Disclosing the Product System Model}

  \input{agg-objective}  % Authors, Readers, review objectives, disclosure constraints

  \input{agg-basic-lca}  % basic Leontief model

  \input{agg-lca-study}  % foreground JLCA model

  \input{agg-disclosure} % disclosure requirements satisfied

\section{Reviewing Product System Models}

Objectives in critical review vary with the study and the reviewer, but they generally include consistency with the ISO 14044 standard, the validity of methods, appropriateness of data used, correctness of interpretations, and transparency of reporting \citep[Section 6.1]{iso14044}.  Although the object of critical review is typically a written report, the use of a structured disclosure of a PSM introduces the possibility that reviewers can inspect the model itself and even verify the computations and reported results.  This section discusses how a disclosure in the format described above can enable robust review of a product system model.  Note that, for purposes of critical review, the objective for data privacy can be omitted because the reviewer may be granted privileged access to confidential data.  We return to the data privacy requirement in the next section.

  \input{agg-foreground} % Af and graph/matrix equivalency

  \input{agg-deps}       % Ad and Bf

\section{Foreground Aggregation}

When an LCA study result is computed, the foreground and background matrices are aggregated into a scalar indicator score (or a small vector of a few scores).  The formulation in Eq.~\ref{eqn:study} and the illustration in Fig.~\ref{fig:foreground} provides a framework for reporting results at varying levels of aggregation.  In this section, I present several different aggregation computations that provide equivalent results for an LCA study.   After that, I show how a foreground containing private data can be partitioned into public and private portions, each of which can be comprehensively checked by a reviewer having privileged access, and subsequently disclosed in a manner that protects the private data.

  \input{agg-agg}         % the elusive Figure 4

  \input{agg-pagg}       % notation for partial aggregation

\section{Discussion}

The current practice of documenting a PSM with a written report sharply limits the ability of readers to interpret and reuse the data.  It also results in the model being frozen to a particular configuration or set of configurations used to generate the published results.  Any review objectives involving sensitivity of the model to parameter variations and/or alternative scenarios are limited to the selections considered by the author.

When a study result is published in an aggregated form, often it is beyond the capacity of the critical reviewer to validate the aggregation result, even if the reviewer has privileged access to confidential materials used in preparing the report.

This paper presents a possible solution by providing a mathematical formalization of the foreground of an LCA study and a functional specification for disclosing the PSM.  A data user or critical reviewer with access to the six disclosure elements, and also adequate access to background datasets and elementary flows used in the model, can reproduce the LCA computation in Eq.~\ref{eqn:study}, and ensure the reported result is correct.  Once equipped with the PSM, the reviewer or data user can potentially go much further.  The graph structure encoded in the foreground matrices can be automatically rendered as a process-flow diagram, supplementing or supplanting the hand-crafted system diagrams that currently accompany ISO reports.  Reviewers or data users who are able to reproduce the PSM from the disclosure can also modify it, adding or removing elements, altering dataset selections, and applying other impact assessment methods.  Moreover, because the disclosure of the PSM includes only references to background data without including the data themselves, it is trivial for item~\ref{itm:bg} to include references to several background databases without adding any additional requirements.

  \input{agg-review}     % review objectives and disclosure constraints

  \input{agg-semantic}   % the need for stable references

\section{Conclusion}

%  \input{publication-foreground}

%  \input{publication-example}

%  \input{publication-discussion}

%  \input{publication-conclusion}

%  \input{publication-acknowledgments}

 \singlespacing
 \bibliography{jabref}
}

\def\supporting{
  \renewcommand{\thefigure}{S\arabic{figure}}
  \renewcommand{\thetable}{S\arabic{table}}

  \maketitle

  \begin{center}
    \Large Supporting Information
  \end{center}

%  \doublespacing

  \input{proofs}

  \input{agg-requirements}

  \input{../publication/publication-example}

%  \input{publication-si-spreadsheet}

%  \input{publication-forms}

%  \input{publication-si-recipe}

%  \input{publication-si-openaccess}

  \singlespacing
  %\bibliography{jabref}
}

\def\unused{
  \input{publication-background}
}

\begin{document}

\mainpaper
%\supporting

%% %\unused

\end{document}