### 2017-11-04 aggregation/agg-abstract.tex
Late afternoon read through

\begin{abstract}
  Many of the challenges facing knowledge synthesis from life cycle assessment (LCA) studies stem from the inability of study authors and readers to formally agree on the structure and content of the product system models used to perform LCA computations.
  This article presents a framework for distinguishing the portions of an LCA study that are attributable to the design of the model from

  isolating the

  computing aggregation results in LCA $\ldots$
\end{abstract}