### 2017-11-16 figures/potato-organic.tex

market for potato, organic [GLO]

{\scriptsize\sffamily
\begin{tabularx}{\textwidth}{|>{\hangindent=3ex}X|c@{~}c@{~}c@{~}c@{~}c@{~}c@{~}c@{~}c@{~}c@{~}|c|}
\hline
(node) Foreground flows -- $A_f$ \rule[-3pt]{0pt}{12pt} & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \\ 
\hline
(0) potato, organic [GLO] (kg) & \refbox  &  &  &  &  &  &  &  &  & \\ 
(1) potato, organic [CH] (kg) & 0.024 & \refbox  &  &  &  &  &  &  &  & \\ 
(2) potato, organic [RoW] (kg) & 0.976 &  & \refbox  &  &  &  &  &  &  & \\ 
(3) potato seed, organic, at farm [CH] (kg) &  &  &  & \refbox  & 0.024 &  &  &  &  & \\ 
(4) potato seed, organic, at farm [GLO] (kg) &  &  &  &  & \refbox  &    1 &    1 &  &  & \\ 
(5) potato seed, organic, for setting [RoW] (kg) &  &  &  &  &  & \refbox  &  &  & 0.976 & \\ 
(6) potato seed, organic, for setting [CH] (kg) &  &  &  &  &  &  & \refbox  &  & 0.024 & \\ 
(7) potato seed, organic, at farm [RoW] (kg) &  &  &  &  & 0.976 &  &  & \refbox  &  & \\ 
(8) potato seed, organic, for setting [GLO] (kg) &  & 0.11 & 0.11 & 0.16 &  &  &  & 0.16 & \refbox  & \\ 
\hline
Foreground Node Weights $\tilde{x}$ &    1 & 0.024 & 0.976 & 0.00315 & 0.131 & 0.128 & 0.00315 & 0.128 & 0.131& \\ 
\hline
\hline
Background Dependencies -- $A_d$ \rule[-3pt]{0pt}{12pt} & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & $\tilde{a_d}$\\ 
\hline
building, multi-storey [GLO] (m3) &  &  &  &  &  & \dependency & \dependency &  &  & 5.24e-06\\ 
electricity, low voltage [GLO] (kWh) &  &  &  &  &  & \dependency &  &  &  & 0.00933\\ 
potato haulm cutting [GLO] (ha) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 5.04e-05\\ 
tillage, hoeing and earthing-up, potatoes [GLO] (ha) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 0.000101\\ 
transport, tractor and trailer, agricultural [GLO] (metric ton*km) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 0.00113\\ 
transport, freight, light commercial vehicle [GLO] (metric ton*km) & \dependency &  &  &  & \dependency &  &  &  &  & 0.0258\\ 
tillage, ploughing [GLO] (ha) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 5.04e-05\\ 
green manure, organic, until March [GLO] (ha) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 5.04e-05\\ 
transport, freight, lorry, unspecified [GLO] (metric ton*km) & \dependency &  &  &  & \dependency &  &  &  &  & 0.404\\ 
transport, freight train [GLO] (metric ton*km) & \dependency &  &  &  & \dependency &  &  &  &  & 0.138\\ 
transport, freight, inland waterways, barge [GLO] (metric ton*km) & \dependency &  &  &  & \dependency &  &  &  &  & 0.0857\\ 
copper oxide [GLO] (kg) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 0.000107\\ 
tillage, harrowing, by spring tine harrow [GLO] (ha) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 5.04e-05\\ 
potato planting [GLO] (ha) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 5.04e-05\\ 
solid manure loading and spreading, by hydraulic loader and spreader [GLO] (kg) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 0.717\\ 
potato grading [GLO] (kg) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  &  1.13\\ 
tillage, harrowing, by rotary harrow [GLO] (ha) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 5.04e-05\\ 
harvesting, by complete harvester, potatoes [GLO] (ha) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 5.04e-05\\ 
transport, freight, sea, transoceanic ship [GLO] (metric ton*km) & \dependency &  &  &  & \dependency &  &  &  &  & 0.525\\ 
liquid manure spreading, by vacuum tanker [GLO] (m3) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 0.000565\\ 
electricity, low voltage [CH] (kWh) &  &  &  &  &  &  & \dependency &  &  & 0.00023\\ 
irrigation [GLO] (m3) &  &  & \dependency &  &  &  &  & \dependency &  & 0.0173\\ 
tillage, currying, by weeder [GLO] (ha) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 0.000101\\ 
application of plant protection product, by field sprayer [GLO] (ha) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 0.000222\\ 
irrigation [CH] (m3) &  & \dependency &  & \dependency &  &  &  &  &  & 0.000426\\ 
\hline
Foreground Emissions -- $B_f$ \rule[-3pt]{0pt}{12pt} & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & $\tilde{b_f}$\\ 
\hline
Input: Occupation, construction site [natural resource, land] (m2*year) &  &  &  &  &  & \dependency & \dependency &  &  & 2.1e-06\\ 
Input: Transformation, from unspecified [natural resource, land] (m2) &  &  &  &  &  & \dependency & \dependency &  &  & 1.05e-06\\ 
Input: Transformation, to industrial area [natural resource, land] (m2) &  &  &  &  &  & \dependency & \dependency &  &  & 1.05e-06\\ 
Input: Energy, gross calorific value, in biomass [natural resource, biotic] (MJ) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  &  3.87\\ 
Output: Phosphate [water, ground-] (kg) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 6.25e-06\\ 
Output: Cadmium, ion [water, ground-] (kg) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 1.14e-09\\ 
Output: Zinc, ion [water, ground-] (kg) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 1.35e-06\\ 
Output: Nitrate [water, ground-] (kg) &  & \dependency & \dependency & \dependency &  &  &  & \dependency &  & 0.00696\\ 
$\ldots$ (31 rows omitted) &  &  &  &  &  &  &  &  & \\ 
\hline
\end{tabularx}
}