### 2018-02-05 aggregation/agg-intro.tex
Streamlined Intro

\section{Introduction}

Life cycle assessment (LCA) is a standard methodology to estimate the potential environmental impacts of products or services by modeling the network of industrial processes that must occur to deliver 
them to a consumer.  The technique is well-established, widely practiced, and draws on an extensive body of standardization and scholarship.
Increasingly, the results of LCA studies are appearing in environmental declarations, corporate sustainability reports, and product marketing information.  LCA is also gaining prominence as a tool for developing and evaluating environmental policy.  
%
At the core of an LCA study is a model of a ``product system,'' which represents a collection of ``processes $\ldots$ performing one or more defined functions'', that describes the life cycle of a product \citep{iso14044}.  The product system model, or PSM, encodes information about how a product is manufactured, distributed, used by the consumer, and what happens to it after it is disposed.  A product system can be divided into a foreground, which denotes the portions of the life cycle whose operations are directly modeled by the study, and a background, which represents the global industrial system \citep{SETAC_inventory_1998}.  While the foreground is often modeled from direct observation or other primary data, background processes are typically drawn from a life cycle inventory (LCI) database prepared by a third party.

Although there has been considerable effort to normalize and harmonize 
LCI database design methodology and data interchange \citep{UNEP_2011, JRC_ILCD_ELCD_2013, Mila_e_Canals_2015, Ingwersen_JLCA_2015}, less attention has been paid to harmonizing PSMs.  These models are often highly complex and include countless modeling decisions, approximations, and assumptions made by a study author \citep{Lloyd2007, reap2008_I}.
Factors such as co-production strategy \citep{Finnveden_1999, Pelletier_2014} can also render study findings unreliable or highly contingent.  Consequently, while comparative evaluations can be made regarding alternative cases or scenarios within a single study, comparisons across studies are much more challenging \citep{Heath2012, Henriksson2014}.  Even when relatively simple and homogeneous systems are considered, wide variations can be found in the results of studies from different authors \citep{van_der_Harst_2013, Turconi_2013}.
A lack of transparency in reporting, particularly regarding definitions of study scope and system boundary,
is a crucial challenge to the interpretation of results in comparative analysis \citep{Cleary2009, Laurent_2014}.  A further challenge to transparency can be found in the use of input data that are confidential or proprietary \citep{Kuczenski_ESD_2017}.  One consequence of all these challenges is the high cost and uneven rigor of critical review, in which the complexities of LCA come face to face with the limitations of current practice \citep{Curran2014}.

As LCA gains prominence, particularly in the policy realm, these problems become more acute.  If LCA is to be used in policy, it will necessarily have the effect of recommending certain product systems, technologies, or approaches at the expense of others; yet in that event it is critically important that there be a consensus among stakeholders that the quantitative results are credible and well-supported \citep{Rainville_2015, McManus_2015}.  As noted above, this level of consensus is hard to realize, in part because of the conflicting implications inherent in different analytic modes.
Plevin et al.\ famously observed in the strongest terms that careless reporting of attributional LCA study results can distort the significance of findings in a policy context \citep{Plevin_2013}, a shortcoming that can be found in other modes as well \citep{Brandao_2014}.

An LCA study result is ultimately an assertion that for some constructed PSM, the delivery of a particular reference flow is associated with a certain amount of environmental impact or potential impact.  But if the PSM itself is ambiguously stated, then the significance of the result is ambiguous as well.  Often the structure and contents of the model are never stated precisely; instead they are \emph{described} in a written report, and the description must be interpreted by a reader.  When a study is reviewed, it is often the written report and \emph{not} the PSM itself that is the object of review.  \cite{iso14071}, which provides guidelines on critical review, emphasizes that review of the PSM or individual datasets is optional.  
%A common theme in all these controversies is the inability for authors and their readers to formally agree on the structure and content of the PSM.
Although the general structure for life cycle assessment can be stated formally, there is no current practice for formally describing the PSM and thus, no automatic mechanisms for interpretation or review of published LCA studies.  Moreover, there is no current practice that permits a reviewer to verify the correctness of the computations inherent in an LCA study if access to the PSM itself is not available.

Because of the diversity of software and techniques for performing LCA, it is important to develop methods for communicating the structure of the PSM that is not dependent on any particular software or database \cite{Kuczenski_JLCA_2018}.    The purpose of this paper is to advance a framework for the formal description of PSM that would allow for sharing, revision, and critical review of models while also enabling authors to protect the privacy of confidential information.  This objective is also in line with the broader imperative throughout industrial ecology to improve transparency and reproducibility of computed results \cite{Hertwich_JIE_2018}.

\endinput

The purpose of this paper is three-fold:
\begin{enumerate}[label=\roman*.]
\item to advance a fram
\item to describe the computation of aggregated LCIA results using a product system model as input data;
\item to show how the framework may be used to protect confidential information while still providing transparency and reproducibility during critical review.
\end{enumerate}

o advance a framework for the formal description of a subset of LCA studies that can be used to document, validate, and reproduce the computations that lead from model specification to LCIA results.  The framework can also

But reproducibility of a computation requires a distinction between the computing algorithm and the input data \citep{Buckheit_1995, }; contemporary LCA lacks this distinction.

What is the point of the new paper?

Some thoughts to put together and rearrange:

 *

 * When a study is reviewed, it is often the written documentation and \emph{not} the product system model itself that is the object of review.

  - In current practice, it is the responsibility of the study author to ensure that the written report accurately describes the model.

  - Because the model itself may not be available for review, it is impossible for a reviewer to directly evaluate the author's claim that the model is accurately represented.

or for distinguishing between the foreground and the background of a model.

 *

 * In the process of aggregating the models, much of the information about their construction is lost.  This is partly by design, in order to protect the privacy of the input data \citep[Ch. 3]{UNEP_2011}.

% \citep{Kuczenski_2017}.

 * However, that same feature is also a shortcoming when considered from the perspective of study transparency, replication, and extension.

 *

 * Because of the nature of the computation, it is straightforward to provide varying degrees of

The aim of this paper is to present a consistent and reproducible formulation of LCA foreground models, and to focus on the aggregation step

 * Documents such as ISO reports, PCRs, or EPDs are uniformly \textit{unstructured}, meaning their interpretation must be secured by a human reading text and graphical images such as system boundary diagrams, but cannot be expressed to a machine.

 *

\oldinput

A common theme in all these controversies is the inability for authors and their readers to formally agree on the structure and content of the product system model.  In the scope of academic research, the transparency of techniques and reproducibility of results are paramount concerns \citep{Mesirov_2010}.  But reproducibility of a computation requires a distinction between the computing algorithm and the input data \citep{Buckheit_1995, Fomel_2009}; contemporary LCA lacks this distinction.  While the ostensible input to an LCIA computation is a product system model, this input cannot be separated from the computational environment used to author it.  Instead, models are distilled into written documents describing LCA results, be they academic papers, ISO reports, PCRs or EPDs.  These are uniformly \textit{unstructured}, meaning their interpretation must be secured by a human reading text and graphical images such as system boundary diagrams, but cannot be expressed to a machine.

The objective of \textit{structured} publication, on the other hand, is to enable the reader of a publication to automatically interpret its contents and, ideally, reproduce its results.  Such a publication is both machine-readable and easily human-interpretable.  In this paper, I propose a framework for the structured publication of product system models in LCA that achieves these dual requirements.

\cruft

%% \textit{Publication}, for the purpose of this article, means disclosing sufficient detail to allow a reader to reproduce an LCA result, or all or part of the model that produced it.  I begin with a set of objectives for structured publication of a study model. Then I show how a simple analytic formulation of study structure can meet these objectives.   I provide examples of structured formulations of studies drawn from the US LCI and Ecoinvent databases.  %% 75 words