### 2018-02-07 aggregation/agg-JIE.tex
Finishing body revision

\documentclass[12pt]{article}

\input{formatting}

%\lhead{\sffamily Disclosure of product system models}
%\rhead{\sffamily Kuczenski -- 2017}

\usepackage{tabularx}
\usepackage[round]{natbib}
\usepackage{siunitx}

\makeatletter
\let\oldsection\section
\renewcommand\section{\@startsection {section}{1}{\z@}%
                                   {-3.5ex \@plus -1ex \@minus -.2ex}%
                                   {2.3ex \@plus.2ex}%
                                   {\normalfont\Large\bfseries <Level 1> }*}
\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
                                     {-3.25ex\@plus -1ex \@minus -.2ex}%
                                     {1.5ex \@plus .2ex}%
                                     {\normalfont\large\bfseries <Level 2> }*}
\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}%
                                     {-3.25ex\@plus -1ex \@minus -.2ex}%
                                     {1.5ex \@plus .2ex}%
                                     {\normalfont\normalsize\bfseries <Level 3> }*}

\makeatother

\usepackage{endnotes}
\let\footnote=\endnote

\bibliographystyle{authordate_bk}

\renewcommand{\topfraction}{0.7}
\renewcommand{\bottomfraction}{0.7}

%\thispagestyle{empty}

\input{psstyle}
\input{../manuscript/draw-matrices}

\title{Disclosure of Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy}
\author{Brandon Kuczenski\\
Institute for Social, Behavioral, and Economic Research\\
University of California, Santa Barbara}
\date{\mydate\today}

\usepackage{lineno}
\let\showgrid\relax
\let\oldtimes\times
\let\times\cdot

\def\mainpaper{

  \maketitle
  \emph{Submitted after revision to the Journal of Industrial Ecology; February 2018}

  \input{agg-abstract}

  \doublespacing
  %\linenumbers

%  {\red http://pubs.acs.org/doi/pdf/10.1021/acs.est.5b01558}

  \input{agg-intro}  % motivations

\section{Disclosing the Product System Model}

  \input{agg-objective}  % Authors, Readers, review objectives, disclosure constraints

  \input{agg-basic-lca}  % basic Leontief model

  \input{agg-lca-study}  % foreground JLCA model

  \input{agg-disclosure} % disclosure requirements satisfied

\section{Reviewing Product System Models}

Objectives in critical review vary with the study and the reviewer, but they generally include consistency with the ISO 14044 standard, the validity of methods, appropriateness of data used, correctness of interpretations, and transparency of reporting \citep[Section 6.1]{iso14044}.  Although the object of critical review is typically a written report, this level of review is not sufficient to ensure the correctness of quantitative results.
The use of a structured disclosure of a PSM introduces the possibility that reviewers can inspect the model itself and even verify the computations and reported results.  This section discusses how a disclosure in the format described above can enable robust review of a product system model. %\footnote{Although ISO 14071 makes explicit that reviewers are not required to review the system model, it is the opinion of both the author and at least one anonymous reviewer that the status quo is insufficient.}
Note that, for purposes of critical review, the objective for data privacy can be omitted because the reviewer may be granted privileged access to confidential data.  We return to the data privacy requirement in the next section.

\input{agg-foreground}

\input{agg-deps}       % Ad and Bf

\section{Foreground Aggregation}

When an LCA study result is computed, the foreground and background matrices are aggregated into a scalar indicator score (or a small vector of a few scores).  The formulation in Eq.~\ref{eqn:study} and the illustration in Fig.~\ref{fig:foreground} provides a framework for reporting results at varying levels of aggregation.  In this section, I present several different aggregation computations that provide equivalent results for an LCA study.   After that, I show how a foreground containing private data can be partitioned into public and private portions, each of which can be comprehensively checked by a reviewer having privileged access, and subsequently disclosed in a manner that protects the private data.

  \input{agg-agg}         % the elusive Figure 4

  \input{agg-pagg}       % notation for partial aggregation

\section{Discussion}

  \input{agg-review}     % review objectives and disclosure constraints

  \input{agg-semantic}   % the need for stable references

%  \input{agg-community}

  \input{agg-conclusion}

%  \input{publication-foreground}

%  \input{publication-example}

%  \input{publication-discussion}

%  \input{publication-conclusion}

%  \input{publication-acknowledgments}

\let\section\oldsection
    \section*{Acknowledgments}

Many thanks to the anonymous reviewers for significant improvement of the manuscript. This work was partially supported by the US National Science Foundation (CCF-1442966).

 \singlespacing

 \theendnotes

 \bibliography{jabref}
}

\def\supporting{
  \setcounter{figure}{0}
  \setcounter{table}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}
  \renewcommand{\thetable}{S\arabic{table}}

  \maketitle

  \begin{center}
    \Large Supporting Information
  \end{center}

%  \doublespacing

  \input{proofs}

  \input{agg-requirements}

  \input{agg-preparing} % Af and graph/matrix equivalency

  \input{agg-multioutput}

  \input{publication-example}

%  \input{publication-si-spreadsheet}

%  \input{publication-forms}

%  \input{publication-si-recipe}

%  \input{publication-si-openaccess}

  \singlespacing
  %\bibliography{jabref}
}

\def\unused{
  \input{publication-background}
}

\begin{document}

\mainpaper
%\supporting

%% %\unused

\end{document}