### 2018-02-07 aggregation/agg-deps.tex

\subsection{Dependencies and Emissions}

The foreground matrix only describes the structure of the PSM; it does not provide any information about inventory or impacts.  The inventory requirements and environmental burdens associated with the model are all described in the dependency matrix $A_d$ and the foreground emission matrix $B_f$.  The dependency matrix shows how foreground nodes depend on activities provided by processes in a background database; the emission matrix shows direct exchanges with the environment.  In both cases, the respective columns of $A_d$ and $B_f$ correspond with columns in $A_f$ and denote dependencies and/or emissions of the corresponding nodes.  If columns of $A_d$ and/or $B_f$ are empty, then the node is simply a pass-through process that serves to connect different foreground elements.  However, if $A_d$ and/or $B_f$ have nonzero entries in that column, then that indicates some transformation process or emission is occurring at that point in the foreground.

Review of the dependencies of foreground nodes must include review of the selection of background processes to which they are connected (i.e. the identities of the columns of $B_x$ referred to in disclosure item~\ref{itm:bg}, the list of background flows). Likewise, the identities of the emissions (entries in disclosure item~\ref{itm:em}) must also be reviewed.